\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{amsmath}

\title{Repechage 2021}
\date{February 2021}

\usepackage{geometry}
\geometry{
 letterpaper,
 left=20mm,
 right=20mm,
 top=20mm,
 bottom=20mm
 }

\setlength{\parindent}{0pt}

\pagestyle{fancy}
\fancyhf{}

\chead{Question 6 \\}
\rfoot{{\thepage} of \pageref{LastPage}}

\begin{document}

Rearrange the equation:
\begin{align*}
    w^2 + 11x^2 -8y^2 -12yz - 10z^2 = 0 \\
    w^2 + 11x^2 = 8y^2 + 12yz + 10z^2. \tag{i}
\end{align*}
Take modulo $4$:
\begin{align*}
    w^2 + 11x^2 &\equiv 8y^2 + 12yz + 10z^2 \pmod 4\\
    w^2 + 3x^2 &\equiv 2z^2 \pmod 4.
\end{align*}
$1$ and $0$ are the only quadratic residues modulo 4, thus: $w^2, x^2, z^2 \equiv 0,1 \pmod 4$ \\
Thus, the only solutions are:
\begin{align}
    w^2 &\equiv 1 \pmod 4,\ x^2 \equiv 1 \pmod 4,\ z^2 \equiv 0 \pmod 4 \\
    w^2 &\equiv 0 \pmod 4,\ x^2 \equiv 0 \pmod 4,\ z^2 \equiv 0 \pmod 4
\end{align}
\\
\textbf{In case (1)}, $w,x$ are odd and $y$ is even. \\
Thus, let $w = 2i + 1$, $x = 2j + 1$, and $z = 2k$. Substitute this into (i) to obtain:
\begin{align*}
    (2i + 1)^2 + 11(2j + 1)^2 &= 8y^2 + 12y(2k) + 10(2k)^2 \\
    4i^2 + 4i + 44j^2 + 44j + 12 &= 8y^2 + 24yk + 40k^2 \\
    i(i + 1) + 11j(j+1) + 3 &= 2y^2 + 6yk + 10k^2.
\end{align*}
Take modulo 2 of the left side to obtain:
$$i(i + 1) + 11j(j+1) + 3 \equiv i(i + 1) + j(j+1) + 1 \pmod 2.$$
Since $i$ and $j$ are integers, $i(i + 1) \equiv j(j+1) \equiv 0 \pmod 2$. Thus:
$$i(i + 1) + j(j+1) + 1 \equiv 1 \pmod 2.$$
Modulo 2 of the right side produces:
$$2y^2 + 6yk + 10k^2 \equiv 0 \pmod 2.$$
This is a contradiction. Thus, case (1) is impossible. \\ \\

\textbf{For case (2)}, $w,x,z$ are all even. Let $w = 2l,\ x = 2m,\ z = 2n$. Substitute this into (i) and simplify:
\begin{align*}
    (2l)^2 + 11(2m)^2 &= 8y^2 + 12y(2n) + 10(2n)^2 \\
    l^2 + 11m^2 &= 2y^2 + 6yn + 10n^2. \tag{ii}
\end{align*}
Now, take modulo $2$ of this equation:
\begin{align*}
    l^2 + 11m^2 &\equiv 2y^2 + 6yn + 10n^2 \pmod 2 \\
    l^2 + m^2 &\equiv 0 \pmod 2.
\end{align*}
The only solutions are:
\begin{align*}
    l^2 \equiv 1 \pmod 2 \ \text{and}\ m^2 \equiv 1 \pmod 2 \tag{2.1} \\
    l^2 \equiv 0 \pmod 2 \ \text{and}\ m^2 \equiv 0 \pmod 2. \tag{2.2}
\end{align*}
\newpage
\textbf{Consider case (2.1)}, $l,m$ must be odd, so let $l = 2p + 1$ and $m = 2q + 1$. Substitute this into (ii) and simplify:
\begin{align*}
    (2p + 1)^2 + 11(2q + 1)^2 &= 2y^2 + 6yn + 10n^2 \\
    2p^2 + 2p + 22q^2 + 22q + 6 &= y^2 + 3yn + 5n^2. \tag{iii}
\end{align*}
The left side is even, so $y$ and $n$ must both be even. Let $y = 2r$, $n = 2s$, substitute into (iii) and simplify:
\begin{align*}
    2p^2 + 2p + 22q^2 + 22q + 6 &= (2r)^2 + 3(2r)(2s) + 5(2s)^2 \\
    2p^2 + 2p + 22q^2 + 22q + 6 &= 4r^2 + 12rs + 20s^2
\end{align*}
Modulo 4 of the right side produces:
$$4r^2 + 12rs + 20s^2 \equiv 0 \pmod 4.$$
Take modulo 4 of the left side to obtain:
\begin{align*}
    2p^2 + 2p + 22q^2 + 22q + 6 &\equiv 2p^2 + 2p + 2q^2 + 2q + 2 \pmod 4 \\
    &\equiv 2p(p + 1) + 2q(q+1) + 2 \pmod 4
\end{align*}

Since $p$ and $q$ are integers, $2p(p + 1) \equiv 2q(q+1) \equiv 0 \pmod 4$. Thus:
$$2p^2 + 2p + 2q^2 + 2q + 2 \equiv 2 \pmod 4.$$
This is a contradiction. Thus, case (2.1) is impossible. \\ \\

\textbf{Consider case (2.2)}, $l$ and $m$ cannot both be zero, as then in (ii), $2y^2 + 6yn + 10n^2 = 0$, thus $l,m,y,n$ are all zero producing $(w,x,y,z) = (0,0,0,0)$ which is excluded. Consider the left side of (ii), by the trivial inequality, $l^2 \ge 0$ and $m^2 \ge 0$ 







must be even, so let $l = 2a$ and $m = 2b$. Again, substitute into (ii) and simplify:
Substitute this into (ii) and simplify:
\begin{align*}
    (2a)^2 + 11(2b)^2 &= 2y^2 + 6yn + 10n^2 \\
    2a^2 + 22b^2 &= y^2 + 3yn + 5n^2. \tag{iv}
\end{align*}
The left side is even, so $y$ and $n$ must both be even. Let $y = 2c$, $n = 2d$, substitute into (iv) and simplify:
\begin{align*}
    2a^2 + 22b^2 &= (2c)^2 + 3(2c)(2d) + 5(2d)^2 \\
    a^2 + 11b^2 &= 2c^2 + 6cd + 10d^2 \\
    \left( \frac{l}{2} \right)^2 + 11\left( \frac{m}{2} \right)^2 &= 2\left( \frac{y}{2} \right)^2 + 6\left( \frac{y}{2} \right)\left( \frac{n}{2} \right) + 10\left( \frac{n}{2} \right)^2
\end{align*}
Thus, if $(a_1,b_1,c_1,d_1)$ is an nonzero integer solution to (ii), then $\left( \frac{a_1}{2}, \frac{b_1}{2}, \frac{c_1}{2}, \frac{d_1}{2} \right)$ must also be such a solution. \\
Now suppose this nonzero integer solution $(a_1,b_1,c_1,d_1)$ exists, then an integer solution $\left( \frac{a_1}{2}, \frac{b_1}{2}, \frac{c_1}{2}, \frac{d_1}{2} \right)$ is a solution, and so is $\left( \frac{a_1}{4}, \frac{b_1}{4}, \frac{c_1}{4}, \frac{d_1}{4} \right)$, $\left( \frac{a_1}{8}, \frac{b_1}{8}, \frac{c_1}{8}, \frac{d_1}{8} \right)$, so on.
Consider $l$ in each solution except the first, that is, $\frac{a_1}{2}, \frac{a_1}{4}, \frac{a_1}{8} \dots$, this is an infinite geometric sequence of distinct integers, call this $S_a$.
Since dividing by $2$ does not change the sign of a number, each term of $S_a$ lies entirely between $a_1$ and $0$. However, since $a_1$ is a finite integer, this interval contains a finite number of distinct integers, while the sequence contains infinite many distinct integers. Thus, such a sequence cannot exist, and thus no solutions to (ii) exist. \\

\textbf{In summary, no solutions to case (1) and case (2) [composed of case (2.1), case (2.2)] exist, so nonzero integer no solutions to (i) exist. Thus, there are no more integer solutions to the original equation.}

\end{document}

$w^2, x^2, z^2 \equiv 0,1 \pmod 4$ as $1$ and $0$ are the quadratic residues of $\pmod 4$.

